The Driving Forces of API Evolution

### 2015-06-03 thesis/paper.tex
etype motivation

Software evolution theory suggests that software programs can be grouped into three different types of systems, namely S-programs, P-programs and E-programs \cite{lehman1980programs}. S-programs are programs which are exactly derivable from a specification. An example of such a program is one which finds the lowest common denominator for two integers. A key characteristic of an S-program is that if its functionality is changed, the end product will be a new product and not an evolved version of the previous program, since the mapping between specification and program is exact. P-programs are programs which are inherently more complex, and ``solve a problem which can be precisely formulated, but whose solution must inevitably reflect an approximation of the real world''. Furthermore, P-programs are programs which are centred around the perception of users, analysts or programmers. This perception is often commonly a mathematical model of the problem. An example of such a program is a program which predicts weather, since it is derived from a set of hydrodynamic equations. E-programs are described as ``inherently even more change prone [than S-programs and P-programs]. They are programs that mechanize a human or societal activity''. These factors can be explained by the difference between P-programs and a E-programs which is that E-programs ``change the very nature of the problem to be solved'' by ``becoming a part of the world it models''. Further, a characteristic of E-programs is that, during development, the specification must be developed by involving predictions of the consequences of the deployment of that system will have, since once the program is deployed ``questions of correctness, appropriateness, and satisfaction arise which inevitably leads to additional pressure for change". Lehman's eight laws of software evolution are said to apply to such E-programs. The laws are presented in Table \ref{table:expl_laws}.

@@

% Given that there is a combination of different interacting elements that do not only concern interfaces to implementations, we would argue that an API can be described as a system \cite{761853,5733835}.

\subsubsection{APIs as Systems}
The API is rather complex, and in addition to the API interface and implementation code, it includes a middle-layer which manages error handling and the access to various 3\textsuperscript{rd} party libraries. In this study, we use the API interface and the middle-layers as units of analysis. Examples of where the responsibilites exceed those of a primitive API interface is that it includes functionality such as error handling and a daemon which facilitates communication between hardware entities. In addition to that, the SDK which includes the API documentation, user manuals and example applications that utilise the API code. Based on these characteristics and the IEEE definition of a system, which reads as follows; ``A combination of interacting elements organised to achieve one or more stated purposes... Organised to accomplish a specific function or set of functions within a specific environment", we consider the API to be a system because of the combination of the interacting parts of the API, coupled with the specific nature of the problem the API is addressing.

% 1. Refer back to background (e-type)
% 2. Make the case for why our API is an e-program

%Specifics about program type etc here
Based on this, we consider APIs to be E-type systems, since they are heavily dependant on the environment and its changes. 
It is therefore not possible to create an exact specification of any non-trivial API.

\section{Methodology} \label{methodology}

%Intro and summary
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis. The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, robson2002real, runeson2009guidelines, yin2013case}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of drivers of API change and that input from the industry is vital to the success of the research.

>>>>

Software evolution theory suggests that software programs can be grouped into three different types of systems, namely S-programs, P-programs and E-programs \cite{lehman1980programs}. S-programs are programs which are exactly derivable from a specification. An example of such a program is one which finds the lowest common denominator for two integers. A key characteristic of an S-program is that if its functionality is changed, the end product will be a new program and not an evolved version of the previous program, since the mapping between specification and program is exact. P-programs are programs which are inherently more complex, and ``solve a problem which can be precisely formulated, but whose solution must inevitably reflect an approximation of the real world''. Furthermore, P-programs are programs which are centred around the perception of users, analysts or programmers. This perception is often commonly a mathematical model of the problem. An example of such a program is a program which predicts weather, since it is derived from a set of hydrodynamic equations. E-programs are described as ``inherently even more change prone [than S-programs and P-programs]. They are programs that mechanize a human or societal activity''. These factors can be explained by the difference between P-programs and a E-programs which is that E-programs ``change the very nature of the problem to be solved'' by ``becoming a part of the world it models''. Further, a characteristic of E-programs is that, during development, the specification must be developed by involving predictions of the consequences of the deployment of that system will have, since once the program is deployed ``questions of correctness, appropriateness, and satisfaction arise which inevitably leads to additional pressure for change". Lehman's eight laws of software evolution are said to apply to such E-programs. The laws are presented in Table \ref{table:expl_laws}.

@@

\subsubsection{APIs as Systems}
The API is rather complex, and in addition to the API interface and implementation code, it includes a middle-layer which manages error handling and the access to various 3\textsuperscript{rd} party libraries. In this study, we use the API interface and the middle-layers as units of analysis. Examples of where the responsibilites exceed those of a primitive API interface is that it includes functionality such as error handling and a daemon which facilitates communication between hardware entities. In addition to that, the SDK which includes the API documentation, user manuals and example applications that utilise the API code. Based on these characteristics and the IEEE definition of a system, which reads as follows; ``A combination of interacting elements organised to achieve one or more stated purposes...Organised to accomplish a specific function or set of functions within a specific environment" \cite{5733835 }, we consider the API to be a system because of the combination of the interacting parts of the API, coupled with the specific nature of the problem the API is addressing.

Building on this, we regard the API examined in this study to be an E-type system based on that it is a part of the world it models. This is evident from that it is a vital part of the cameras' and the only option for developing software for the cameras. The API is also by company A considered to be a vital part of the cameras' software ecosystem. This factor also corresponds with the characteristic of E-type systems which relates to the system being susceptible to change from pressure that occurs after the system has been deployed. Similarly, we do not consider the API to be a P-type system, based on that the system is virtually impossible to model because of its high dependenace with the environment it operates in.

\section{Methodology} \label{methodology}
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis. The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, robson2002real, runeson2009guidelines, yin2013case}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of drivers of API change and that input from the industry is vital to the success of the research.
